Manuscript

### 2014-10-11 inst/simulation/manuscript.Rmd
...

context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
scenarios <- c("A","B","C","D","E","F")
modelNames <- c(paste("SmoothModel", "nbinomial", "matern", "ar1", sep = "-"), paste("SmoothModel", "nbinomial", "ar1", sep = "-"), "FMPModel")
iterations <- as.integer(c(1,1,1,1,3,1))
populationSizeOverEstimate <- 1e100

@@

```{r figure-3-test, eval=FALSE}
study <- getStudy(scenario="A")
validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
x <- validation$validateTemporalPopulationSize(modelNames[1])
validation$summarizePopulationSize(x)
```

```{r figure-3-init-200}
study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
populationSizes <- validation$getValidationTemporalPopulationSizes(modelNames=modelNames)
validation$summarizePopulationSize(populationSizes)

meanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
  data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
})
logPopulationSizes <- populationSizes
logPopulationSizes$Observed <- logPopulationSizes$Observed + 1
logPopulationSizes$Estimated <- logPopulationSizes$Estimated + 1
logMeanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
  data.frame(True=mean(x$Observed+1), Estimated=mean(x$Estimated+1), Scenario=x$scenario)
})
#logPopulationSizes$Observed <- log10(logPopulationSizes$Observed + 1)
#logPopulationSizes$Estimated <- log10(logPopulationSizes$Estimated + 1)
#logMeanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
#  data.frame(True=mean(log10(x$Observed+1)), Estimated=mean(log10(x$Estimated+1)), Scenario=x$scenario)
#})
```

@@

```{r figure-3b}
plotLogPopulationSize <- function(size, meanSize) {
  p <- ggplot(size, aes(Observed, Estimated, colour=modelName)) +
    #geom_point(size=1) +
    #stat_smooth(method="lm", aes(colour=modelName), se=F) +
    geom_point(size=1, colour="darkgray") +
    stat_smooth(method="lm", se=F, colour="black") +
    #scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
    #scale_colour_manual("Model", values=c("darkgray","black")) +
    geom_abline(colour="black", linetype="longdash") + 
    #geom_point(data=meanSize, aes(True, Estimated, colour=modelName), size=2, shape=15) +
    geom_point(data=meanSize, aes(True, Estimated), size=2, shape=15, colour="black") +
    facet_wrap(~Scenario, scales="free") +
    scale_x_log10() + scale_y_log10() +
    xlab("True population size") + ylab("Estimated population size") +
    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"))
  print(p)
  return(p)
}

p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), subset(logMeanPopulationSizes, modelName==modelNames[1]))
saveFigure(p, "manuscript-figure-3b1.png", width=8, height=4)
p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[2]), subset(logMeanPopulationSizes, modelName==modelNames[2]))
saveFigure(p, "manuscript-figure-3b2.png", width=8, height=4)
p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), subset(logMeanPopulationSizes, modelName==modelNames[3]))
saveFigure(p, "manuscript-figure-3b3.png", width=8, height=4)
```

@@

meanPopulationSizes2000 <- ddply(populationSizes2000, .(scenario, modelName), function(x) {
  data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
})
logPopulationSizes2000 <- populationSizes2000
logPopulationSizes2000$Observed <- (logPopulationSizes2000$Observed + 1)
logPopulationSizes2000$Estimated <- (logPopulationSizes2000$Estimated + 1)
logMeanPopulationSizes2000 <- ddply(populationSizes2000, .(scenario, modelName), function(x) {
  data.frame(True=mean((x$Observed+1)), Estimated=mean((x$Estimated+1)), Scenario=x$scenario)
})
```

@@

meanPopulationSizes10days <- ddply(populationSizes10days, .(scenario, modelName), function(x) {
  data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
})
logPopulationSizes10days <- populationSizes10days
logPopulationSizes10days$Observed <- (logPopulationSizes10days$Observed + 1)
logPopulationSizes10days$Estimated <- (logPopulationSizes10days$Estimated + 1)
logMeanPopulationSizes10days <- ddply(populationSizes10days, .(scenario, modelName), function(x) {
  data.frame(True=mean((x$Observed+1)), Estimated=mean((x$Estimated+1)), Scenario=x$scenario)
})
```

@@

```{r table-1abc}
prettyModelName <- function(x) {
  lookup <- data.frame(x=c("SmoothModel-nbinomial-matern-ar1", "SmoothModel-nbinomial-ar1", "FMPModel") , y=c("ST", "T", "I"), stringsAsFactors=F)
  lookup$y[match(x, lookup$x)]
}

getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, full=F, file) {
  library(plyr)
  x <- ddply(populationSizes, .(Scenario, modelName), function(x) {
    y <- lm(Estimated~Observed, x)
    data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2))
  })
  y <- unique(meanPopulationSizes)
  y$Scenario <- y$scenario
  y$EstDev <- y$Estimated / y$True
  z <- plyr::join(x, y)
  z$Model <- prettyModelName(z$modelName)
  z$Scenario <- substr(z$Scenario, 1, 1)
  if (full) return(z)

dir <- normalizePath(context$figuresDirectory)
getPopulationSizesTable(populationSizes, meanPopulationSizes, file=file.path(dir, "table-1a.csv"))
getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, file=file.path(dir, "table-1b.csv"))
getPopulationSizesTable(populationSizes10days, meanPopulationSizes10days, file=file.path(dir, "table-1c.csv"))
```

>>>>

context <- Context$new(resultDataDirectory=wd.data.results, processedDataDirectory=wd.data.processed, rawDataDirectory=wd.data.raw, scratchDirectory=wd.scratch, figuresDirectory=wd.figures)
scenarios <- c("A","B","C","D","E","F")
modelNames <- c(paste("SmoothModel", "nbinomial", "matern", "ar1", sep = "-"), paste("SmoothModel", "nbinomial", "ar1", sep = "-"), "FMPModel")
iterations <- as.integer(c(1,1,1,1,3,1))
populationSizeOverEstimate <- 1e999

prettyModelName <- function(x) {
  lookup <- data.frame(x=c("SmoothModel-nbinomial-matern-ar1", "SmoothModel-nbinomial-ar1", "FMPModel") , y=c("ST", "T", "I"), stringsAsFactors=F)
  lookup$y[match(x, lookup$x)]
}

@@

```{r figure-3-init}
getPopulationSizes <- function(scenarios, modelNames, populationSizeOverEstimate=Inf) {
  study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
  validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
  populationSizes <- validation$getValidationTemporalPopulationSizes(scenarios=scenarios, modelNames=modelNames)
  #validation$summarizePopulationSize(populationSizes)  
}

getMeanPopulationSizes <- function(populationSizes) {
  meanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
    data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
  })
  return(meanPopulationSizes)
}

getLogPopulationSizes <- function(populationSizes) {
  logPopulationSizes <- populationSizes
  logPopulationSizes$Observed <- logPopulationSizes$Observed + 1
  logPopulationSizes$Estimated <- logPopulationSizes$Estimated + 1
  return(logPopulationSizes)
}

getLogMeanPopulationSizes <- function(populationSizes) {  
  logMeanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
    data.frame(True=10^mean(log10(x$Observed+1)), Estimated=10^mean(log10(x$Estimated+1)), Scenario=x$scenario)
  })
  return(logMeanPopulationSizes)
}

plotLogPopulationSize <- function(size, meanSize, modelName, xlim, ylim) {
  library(ggplot2)
  library(scales)
  p <- ggplot(size, aes(Observed, Estimated, colour=modelName)) +
    #geom_point(size=1) +
    #stat_smooth(method="lm", aes(colour=modelName), se=F) +
    geom_point(size=1, colour="darkgray") +
    stat_smooth(method="lm", se=F, colour="black") +
    #scale_colour_manual("Model", values=c("lightgray", "darkgray","black"), labels=c("I", "T", "ST")) +
    #scale_colour_manual("Model", values=c("darkgray","black")) +
    geom_abline(colour="black", linetype="longdash") + 
    #geom_point(data=meanSize, aes(True, Estimated, colour=modelName), size=2, shape=15) +
    geom_point(data=meanSize, aes(True, Estimated), size=2, shape=15, colour="black") +
    facet_wrap(~Scenario, scales="free") +
    coord_cartesian(xlim=xlim, ylim=ylim) +
    scale_x_log10() + scale_y_log10() +
    xlab("True population size") + ylab("Estimated population size") + ggtitle(paste("Model: ", prettyModelName(modelName[1]))) +
    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=12))
  print(p)
  return(p)
}
```

```{r figure-3-init}
populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames)
populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames)
populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames)

meanPopulationSizes  <- getMeanPopulationSizes(populationSizes)
logPopulationSizes <- getLogPopulationSizes(populationSizes)
logMeanPopulationSizes  <- getLogMeanPopulationSizes(populationSizes)

meanPopulationSizes2000  <- getMeanPopulationSizes(populationSizes2000)
logPopulationSizes2000 <- getLogPopulationSizes(populationSizes2000)
logMeanPopulationSize2000  <- getLogMeanPopulationSizes(populationSizes2000)

meanPopulationSizes10Days  <- getMeanPopulationSizes(populationSizes10Days)
logPopulationSizes10Days <- getLogPopulationSizes(populationSizes10Days)
logMeanPopulationSizes10Days  <- getLogMeanPopulationSizes(populationSizes10Days)
#validation$summarizePopulationSize(populationSizes10Days)
```

```{r figure-3abc}
p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[1]), subset(logMeanPopulationSizes, modelName==modelNames[1]), modelName=modelNames[1], xlim=c(1,10000), ylim=c(1,10000))
saveFigure(p, "manuscript-figure-3a1.png", width=8, height=4)
p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[2]), subset(logMeanPopulationSizes, modelName==modelNames[2]), modelName=modelNames[2], xlim=c(1,10000), ylim=c(1,10000))
saveFigure(p, "manuscript-figure-3a2.png", width=8, height=4)
p <- plotLogPopulationSize(subset(logPopulationSizes, modelName==modelNames[3]), subset(logMeanPopulationSizes, modelName==modelNames[3]), modelName=modelNames[3], xlim=c(1,10000), ylim=c(1,10000))
saveFigure(p, "manuscript-figure-3a3.png", width=8, height=4)

xlim <- c(100,10000)
ylim <- xlim
p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[1]), subset(logMeanPopulationSizes2000, modelName==modelNames[1]), modelName=modelNames[1], xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3b1.png", width=8, height=4)
p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[2]), subset(logMeanPopulationSizes2000, modelName==modelNames[2]), modelName=modelNames[2], xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3b2.png", width=8, height=4)
p <- plotLogPopulationSize(subset(logPopulationSizes2000, modelName==modelNames[3]), subset(logMeanPopulationSizes2000, modelName==modelNames[3]), modelName=modelNames[3], xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3b3.png", width=8, height=4)

p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[1]), subset(logMeanPopulationSizes10Days, modelName==modelNames[1]), modelName=modelNames[1], xlim=c(1,10000), ylim=c(1,10000))
saveFigure(p, "manuscript-figure-3c1.png", width=8, height=4)
p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[2]), subset(logMeanPopulationSizes10Days, modelName==modelNames[2]), modelName=modelNames[2], xlim=c(1,10000), ylim=c(1,10000))
saveFigure(p, "manuscript-figure-3c2.png", width=8, height=4)
p <- plotLogPopulationSize(subset(logPopulationSizes10Days, modelName==modelNames[3]), subset(logMeanPopulationSizes10Days, modelName==modelNames[3]), modelName=modelNames[3], xlim=c(1,10000), ylim=c(1,10000))
saveFigure(p, "manuscript-figure-3c3.png", width=8, height=4)
```

```{r table-1abc}
getPopulationSizesTable <- function(populationSizes, meanPopulationSizes, full=F, file) {
  library(plyr)
  x <- ddply(populationSizes, .(Scenario, modelName), function(x) {
    y <- lm(Estimated~Observed, x)
    data.frame(Intercept=round(coef(y)[1], 0), Slope=round(coef(y)[2], 2), R2=round(summary(y)$r.squared, 2))
  })
  y <- unique(meanPopulationSizes)
  y$Scenario <- y$scenario
  y$EstDev <- y$Estimated / y$True
  z <- plyr::join(x, y)
  z$Model <- prettyModelName(z$modelName)
  z$Scenario <- substr(z$Scenario, 1, 1)
  if (full) return(z)

dir <- normalizePath(context$figuresDirectory)
getPopulationSizesTable(populationSizes, meanPopulationSizes, file=file.path(dir, "table-1a.csv"))
getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, file=file.path(dir, "table-1b.csv"))
getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, file=file.path(dir, "table-1c.csv"))
```

```{r, eval=FALSE}
study <- getStudy(scenario="A") # Get study object, scenario doesn't matter
validation <- Validation(study=study, populationSizeOverEstimate=populationSizeOverEstimate)
populationSizes <- validation$getValidationTemporalPopulationSizes(modelNames=modelNames)
validation$summarizePopulationSize(populationSizes)

meanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
  data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
})
logPopulationSizes <- populationSizes
logPopulationSizes$Observed <- logPopulationSizes$Observed + 1
logPopulationSizes$Estimated <- logPopulationSizes$Estimated + 1
logMeanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
  data.frame(True=10^mean(log10(x$Observed+1)), Estimated=10^mean(log10(x$Estimated+1)), Scenario=x$scenario)
})
#logPopulationSizes$Observed <- log10(logPopulationSizes$Observed + 1)
#logPopulationSizes$Estimated <- log10(logPopulationSizes$Estimated + 1)
#logMeanPopulationSizes <- ddply(populationSizes, .(scenario, modelName), function(x) {
#  data.frame(True=mean(log10(x$Observed+1)), Estimated=mean(log10(x$Estimated+1)), Scenario=x$scenario)
#})
```

@@

@@

meanPopulationSizes2000 <- ddply(populationSizes2000, .(scenario, modelName), function(x) {
  data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
})
logPopulationSizes2000 <- populationSizes2000
logPopulationSizes2000$Observed <- (logPopulationSizes2000$Observed + 1)
logPopulationSizes2000$Estimated <- (logPopulationSizes2000$Estimated + 1)
logMeanPopulationSizes2000 <- ddply(populationSizes2000, .(scenario, modelName), function(x) {
  data.frame(True=10^mean(log10(x$Observed+1)), Estimated=10^mean(log10(x$Estimated+1)), Scenario=x$scenario)
})
```

```{r figure-3c123, eval=FALSE}
```

@@

meanPopulationSizes10days <- ddply(populationSizes10days, .(scenario, modelName), function(x) {
  data.frame(True=mean(x$Observed), Estimated=mean(x$Estimated), Scenario=x$scenario)
})
logPopulationSizes10days <- populationSizes10days
logPopulationSizes10days$Observed <- (logPopulationSizes10days$Observed + 1)
logPopulationSizes10days$Estimated <- (logPopulationSizes10days$Estimated + 1)
logMeanPopulationSizes10days <- ddply(populationSizes10days, .(scenario, modelName), function(x) {
  data.frame(True=10^mean(log10(x$Observed+1)), Estimated=10^mean(log10(x$Estimated+1)), Scenario=x$scenario)
})
```

```{r figure-3d123, eval=FALSE}
```

@@
